---
output: html_document
---

<h1><center> Chicago Crime Report by Remya Kannan </center></h1>

<br>


<h4> Crime in Chicago</h4>
<p> Crime in Chicago has been tracked by the Chicago Police Department's Bureau of Records since the beginning of the 20th century.Chicago was responsible for nearly half of 2016's increase in homicides in the US.As of 2017, Chicago's homicide rate is significantly higher when compared to the larger American cities of New York and Los Angeles, but lower when compared to smaller American cities.</p> 



<br>
<h4>The importance of the dataset</h4>

1. Chicago Crime Numbers have been in the media for all the wrong reasons. 
2. President widely used these numbers during his presidential campaign
3. He still talks about it.
4. These numbers are impacting the overall numbers of the country.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.align = "center")

library(highcharter)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(plotly)
library(lubridate)
library(xts)
library(maps)
library(ggmap)
library(gridExtra)

setwd("/Users/remya/Desktop/Depaul_5th_quarter/CSC_465/Project")

chicagocrimes20122016 <- read.csv("Chicago_Crimes_2012_to_2017.csv")

## Reading the data
chicagocrimes20122016 <- chicagocrimes20122016[chicagocrimes20122016$Year!='2017',]


### Working with Date and Time
chicagocrimes20122016$Day <- factor(day(as.POSIXlt(chicagocrimes20122016$Date, format="%m/%d/%Y %I:%M:%S %p")))
chicagocrimes20122016$Month <- factor(month(as.POSIXlt(chicagocrimes20122016$Date, format="%m/%d/%Y %I:%M:%S %p"), label = TRUE))
chicagocrimes20122016$Year <- factor(year(as.POSIXlt(chicagocrimes20122016$Date, format="%m/%d/%Y %I:%M:%S %p")))
chicagocrimes20122016$Weekday <- factor(wday(as.POSIXlt(chicagocrimes20122016$Date, format="%m/%d/%Y %I:%M:%S %p"), label = TRUE))

chicagocrimes20122016$Date <- as.Date(chicagocrimes20122016$Date, "%m/%d/%Y %I:%M:%S %p")

## Creating timeseries
by_Date <- na.omit(chicagocrimes20122016) %>% group_by(Date) %>% summarise(Total = n())
tseries <- xts(by_Date$Total, order.by=as.POSIXct(by_Date$Date))


## Creating timeseries of arrests made
Arrests_by_Date <- na.omit(chicagocrimes20122016[chicagocrimes20122016$Arrest == 'True',]) %>% group_by(Date) %>% summarise(Total = n())
arrests_tseries <- xts(Arrests_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

## BY Location
by_location <- chicagocrimes20122016 %>% group_by(Location.Description) %>% summarise(Total = n()) %>% arrange(desc(Total))

## By Primary Type
by_type <- chicagocrimes20122016 %>% group_by(Primary.Type) %>% summarise(Total = n()) %>% arrange(desc(Total))
  

## By District
by_district <- chicagocrimes20122016 %>% group_by(District) %>% summarise(Total = n()) %>% arrange(desc(Total))


## By Primary Type
by_ward <- chicagocrimes20122016 %>% group_by(Ward) %>% summarise(Total = n()) %>% arrange(desc(Total))


## By FBI Code
by_fbi <- chicagocrimes20122016 %>% group_by(FBI.Code) %>% summarise(Total = n()) %>% arrange(desc(Total))

## By Arrest
by_arrest <- chicagocrimes20122016 %>% group_by(Arrest) %>% summarise(Total = n()) %>% arrange(desc(Total))

## By Domestic
by_domestic <- chicagocrimes20122016 %>% group_by(Domestic) %>% summarise(Total = n()) %>% arrange(desc(Total))

## By Year
by_year <- chicagocrimes20122016 %>% group_by(Year) %>% summarise(Total = n()) %>% arrange(desc(Total))

## Lat and Long of Crimes
LatLonCounts <- as.data.frame(table(round(chicagocrimes20122016$Longitude,2), round(chicagocrimes20122016$Latitude,2)))
LatLonCounts$Long <- as.numeric(as.character(LatLonCounts$Var1))
LatLonCounts$Lat <- as.numeric(as.character(LatLonCounts$Var2))
LatLonCounts2 <- subset(LatLonCounts, Freq > 0)

## Lat and Long of Arrests
arrests_data <- na.omit(chicagocrimes20122016[chicagocrimes20122016$Arrest == 'True',])
LatLonArrestsCounts <- as.data.frame(table(round(arrests_data$Longitude,2), round(arrests_data$Latitude,2)))
LatLonArrestsCounts$Long <- as.numeric(as.character(LatLonArrestsCounts$Var1))
LatLonArrestsCounts$Lat <- as.numeric(as.character(LatLonArrestsCounts$Var2))
LatLonArrestsCounts2 <- subset(LatLonArrestsCounts, Freq > 0)

```


```{r}

hchart(tseries, name = "Crimes") %>% 
  hc_add_series(arrests_tseries, name = "Arrests") %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>%
  hc_title(text = "Times Series plot of Chicago Crimes and Arrests") %>%
  hc_legend(enabled = TRUE)

```


<br>
<br>

From the above time series plot, we observe that the crime rate in Chicago has decreased in 2016 when compared to 2012. There is a clear indication that crime numbers oncrease during the middle of the year mostly during the summer months and drops towards the end/starting of the year (winter months). Another interesting note is the arrest rate is lower compared to the number of crimes. Additionally, crimes distibuted over months would be interesting to look into.

<br>
<br>


```{r}

hchart(arrests_tseries) %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_title(text = "Times Series plot of Arrests made in Chicago (2012-2016)") %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px"))

```

<br>
<br>
 
The plot above shows the decrease in the arrest rate from 2012-2016! 

<br>
<br>


```{r fig.height=7, fig.width=9}


arrests_count <- arrests_data %>% group_by(Year, Month) %>% summarise(Total = n())

arrests <- ggplot(arrests_count, aes(Year, Month, fill = Total)) +
  geom_tile(size = 1, color = "white") +
  scale_fill_viridis()  +
  geom_text(aes(label=Total), color='white') +
  ggtitle("Arrests by Year and Month(2012-2016)")


crime_count <- chicagocrimes20122016 %>% group_by(Year, Month) %>% summarise(Total = n())

crimes <- ggplot(crime_count, aes(Year, Month, fill = Total)) +
  geom_tile(size = 1, color = "white") +
  scale_fill_viridis()  +
  geom_text(aes(label=Total), color='white') +
  ggtitle("Crimes by Year and Month(2012-2016)")


grid.arrange(crimes, arrests, ncol = 2)

```


<br>

The heatmap indicates the decrease in the number of arrests by more than half between 2012 and 2016 but the crimes have not reduced at the same rate. 
<br>

<!-- ```{r} -->

<!-- hchart(by_year, "column", hcaes(x = Year, y = Total, color = Year)) %>% -->
<!--   hc_colorAxis(stops = color_stops(n = 10, colors = c("#440154", "#21908C", "#FDE725"))) %>% -->
<!--   hc_add_theme(hc_theme_darkunica()) %>% -->
<!--   hc_title(text = "Number of Crimes by Year") %>% -->
<!--   hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>% -->
<!--   hc_legend(enabled = FALSE) -->

<!-- ``` -->

<br>
<br>


```{r}

hchart(by_location[1:20,], "column", hcaes(x = Location.Description, y = Total, color = Total)) %>%
  hc_colorAxis(stops = color_stops(n = 10, colors = c("#440154", "#21908C", "#FDE725"))) %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_title(text = "Locations with most Crimes - Top 20") %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>%
  hc_legend(enabled = FALSE)

```

<br>
<br>

From the above time plot we observe that the streets are the most common location for crimes to occur followed by apartments and residence.

<br>
<br>

#### Let us understand how crime numbers vary in some of these top crime locations 

<br>
<br>

```{r}

streets <- chicagocrimes20122016[chicagocrimes20122016$Location.Description=="STREET",]
## Creating timeseries
streets_by_Date <- na.omit(streets) %>% group_by(Date) %>% summarise(Total = n())
streets_tseries <- xts(streets_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

residence <- chicagocrimes20122016[chicagocrimes20122016$Location.Description=="RESIDENCE",]
## Creating timeseries
residence_by_Date <- na.omit(residence) %>% group_by(Date) %>% summarise(Total = n())
residence_tseries <- xts(residence_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

apartment <- chicagocrimes20122016[chicagocrimes20122016$Location.Description=="APARTMENT",]
## Creating timeseries
apartment_by_Date <- na.omit(apartment) %>% group_by(Date) %>% summarise(Total = n())
apartment_tseries <- xts(apartment_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

sidewalk <- chicagocrimes20122016[chicagocrimes20122016$Location.Description=="SIDEWALK",] 
## Creating timeseries
sidewalk_by_Date <- na.omit(sidewalk) %>% group_by(Date) %>% summarise(Total = n())
sidewalk_tseries <- xts(sidewalk_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

hchart(streets_tseries, name = "Streets") %>% 
  hc_add_series(residence_tseries, name = "Residence") %>% 
  hc_add_series(apartment_tseries, name = "Apartment") %>%
  hc_add_series(sidewalk_tseries, name = "Sidewalk") %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>%
  hc_title(text = "Crimes in Streets/Residence/Apartment/Sidewalk") %>%
  hc_legend(enabled = TRUE)

```


<br>
<br>

The number of crimes have significantly reduced in the top crime locations.

<br>
<br>


```{r}

hchart(by_type, "column", hcaes(Primary.Type, y = Total, color = Total)) %>%
  hc_colorAxis(stops = color_stops(n = 10, colors = c("#440154", "#21908C", "#FDE725"))) %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_title(text = "Crime Types") %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>%
  hc_legend(enabled = FALSE)

```


<br>
<br>

Taking a look at the popular crimes, theft is the topmost followed by battery, narcotics and criminal damage.

<br>
<br>

#### Let us see how crime numbers vary in top crimes

<br>
<br>

```{r}

thefts <- chicagocrimes20122016[chicagocrimes20122016$Primary.Type=="THEFT",] 
## Creating timeseries
thefts_by_Date <- na.omit(thefts) %>% group_by(Date) %>% summarise(Total = n())
thefts_tseries <- xts(thefts_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

battery <- chicagocrimes20122016[chicagocrimes20122016$Primary.Type=="BATTERY",] 
## Creating timeseries
battery_by_Date <- na.omit(battery) %>% group_by(Date) %>% summarise(Total = n())
battery_tseries <- xts(battery_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

criminals <- chicagocrimes20122016[chicagocrimes20122016$Primary.Type=="CRIMINAL DAMAGE",]
## Creating timeseries
criminals_by_Date <- na.omit(criminals) %>% group_by(Date) %>% summarise(Total = n())
criminals_tseries <- xts(criminals_by_Date$Total, order.by=as.POSIXct(by_Date$Date))

narcotics <- chicagocrimes20122016[chicagocrimes20122016$Primary.Type=="NARCOTICS",] 
## Creating timeseries
narcotics_by_Date <- na.omit(narcotics) %>% group_by(Date) %>% summarise(Total = n())
narcotics_tseries <- xts(narcotics_by_Date$Total, order.by=as.POSIXct(by_Date$Date))


hchart(thefts_tseries, name = "Thefts") %>% 
  hc_add_series(battery_tseries, name = "Battery") %>% 
  hc_add_series(criminals_tseries, name = "Criminal Damage") %>%
  hc_add_series(narcotics_tseries, name = "Narcotics") %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px")) %>%
  hc_title(text = "Crimes in Thefts/Battery/Criminal Damage/Narcotics") %>%
  hc_legend(enabled = TRUE)


```


<br>
<br>

From the above pllot we observe that the number of narcotics crimes have reduced and the number of thefts and battery crimes have remained the same.

<br>
<br>

### How big is the increase in homicides in Chicago?


```{r}

homicide <- chicagocrimes20122016[chicagocrimes20122016$Primary.Type=="HOMICIDE",] 

homicide_year <-  homicide %>% group_by(Year) %>% summarise(Total = n())

hchart(homicide_year, "column", hcaes(Year, Total, color = Year)) %>%
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_title(text = "Homicide 2012-2016")  %>%
  hc_credits(enabled = TRUE, text = "Sources: City of Chicago Administration and the Chicago Police Department", style = list(fontSize = "12px"))

```

<br>
<br>

```{r}


homicide_count <- homicide %>% group_by(Year, Month) %>% summarise(Total = n())

ggplot(homicide_count, aes(Year, Month, fill = Total)) +
  geom_tile(size = 1, color = "white") +
  scale_fill_viridis()  +
  geom_text(aes(label=Total), color='white') +
  ggtitle("Homicides in Chicago (2012-2016)")

```


<br>
<br>

There is a significant increase in the number of homicides in Chicago in 2016 compared to previous years. However, there is very little development on the reasons for the increase in the number. Being one of the biggest cities, Chicago does have an impact on the overall crime numbers of the country. Hence the talk about crime even though a lot of other smaller cities have bigger crime rates.



