---
title: "Final Project"
author: "Remya Kannan"
date: "June 10, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

### Step 1 : Load the libraries
```{r}
# load RSiena commands:
library(RSiena)
library(parallel)
```

### Step 2 : Set the path 
```{r}
# set working directory to where the data are:
setwd("/Users/Vinayak/Desktop/Remya/Quarter 3/Social Network Analysis/Project Material")

# Read data sets:
load("Glasgow-friendship.RData")  # friendship networks
load("Glasgow-demographic.RData") # for gender data
load("Glasgow-substances.RData")  # for alcoholol data
load("Glasgow-lifestyle.RData")   # for music listening data
load("Glasgow-selections.RData")  # for 129 respondents used in SSW
```

### Step 3 : Basic definitions in SIENA for the behavioural attributes
```{r}
rockscale <- matrix(nrow=length(age),ncol=3)
rownames(rockscale) <- rownames(music1)
rockscale[,1] <- rowSums(music1[,colnames(music1) %in% c("rock","indie","heavy","grunge")])
rockscale[,2] <- rowSums(music2[,colnames(music2) %in% c("rock","indie","heavy","grunge")])
rockscale[,3] <- rowSums(music3[,colnames(music3) %in% c("rock","indie","heavy","grunge")])
elitescale <- matrix(nrow=length(age),ncol=3)
rownames(elitescale) <- rownames(music1)
elitescale[,1] <- rowSums(music1[,colnames(music1) %in% c("folk","jazz","classical")])
elitescale[,2] <- rowSums(music2[,colnames(music2) %in% c("folk","jazz","classical")])
elitescale[,3] <- rowSums(music3[,colnames(music3) %in% c("folk","jazz","classical")])
chartscale <- matrix(nrow=length(age),ncol=3)
rownames(chartscale) <- rownames(music1)
chartscale[,1] <- rowSums(music1[,colnames(music1) %in% c("techno","chart","dance","rave")])
chartscale[,2] <- rowSums(music2[,colnames(music2) %in% c("techno","chart","dance","rave")])
chartscale[,3] <- rowSums(music3[,colnames(music3) %in% c("techno","chart","dance","rave")])
```

### Step 4 : Convert valued friendship to binary value
```{r}
friendship.1[friendship.1==2] <- 1
friendship.2[friendship.2==2] <- 1
friendship.3[friendship.3==2] <- 1
```

### Step 5 : Identify dependent network and behaviour variables and the constant covariate
```{r}
# Identify dependent network variable:
friendship <- sienaNet(array(c(friendship.1[selection129,selection129],
                               friendship.2[selection129,selection129],friendship.3[selection129,selection129]),
                             dim=c(129,129,3)))

# Identify dependent behavior variables:
drinking <- sienaNet(alcohol[selection129,],type="behavior")
rock <- sienaNet(rockscale[selection129,],type="behavior")
elite <- sienaNet(elitescale[selection129,],type="behavior")
chart <- sienaNet(chartscale[selection129,],type="behavior")

# Identify constant covariate:
sexF <- coCovar(sex.F[selection129])
```

### Step 6 : SIENA model 
```{r}
# Bind data together for Siena analysis:
SSWdata <- sienaDataCreate(friendship,drinking,rock,elite,chart,sexF)

# Create effects object for model specification:
SSWeffects <- getEffects(SSWdata)

#Give some estimation settings.
#Here 4 phase 2 subphases and 1000 phase 3 iterations are used.
# These settings are the standard and work well for most datasets.
estimationSettings <- sienaAlgorithmCreate(projname='SSW-results')

# Specify the model:
SSWeffects <- includeEffects(SSWeffects,nbrDist2)
SSWeffects <- includeEffects(SSWeffects,egoX,altX,sameX,interaction1="sexF")
SSWeffects <- includeEffects(SSWeffects,egoX,altX,simX,interaction1="drinking")
SSWeffects <- includeEffects(SSWeffects,egoX,altX,simX,interaction1="rock")
SSWeffects <- includeEffects(SSWeffects,egoX,altX,simX,interaction1="elite")
SSWeffects <- includeEffects(SSWeffects,egoX,altX,simX,interaction1="chart")
SSWeffects <- includeEffects(SSWeffects,name="drinking",totSim,interaction1="friendship")
SSWeffects <- includeEffects(SSWeffects,name="rock",totSim,interaction1="friendship")
SSWeffects <- includeEffects(SSWeffects,name="elite",totSim,interaction1="friendship")
SSWeffects <- includeEffects(SSWeffects,name="chart",totSim,interaction1="friendship")

# Shorthand to include all effects from covariates/behaviors on behaviors:
SSWeffects$include[SSWeffects$shortName=="effFrom" & SSWeffects$type=="eval"] <- TRUE
# de-select quadratric shape parameters (included by default):
SSWeffects$include[SSWeffects$shortName=="quad"] <- FALSE
# check how model specification looks like:
SSWeffects

# Now create a model object:
SSWmodel <- sienaModelCreate(useStdInits=TRUE,projname='SSW-results')

# Estimate the model (using two cores of the processor to speed up the process):
SSWresults <- siena07(SSWmodel,data=SSWdata,effects=SSWeffects,useCluster=TRUE,
                      initC=TRUE,nbrNodes=2)


# Take a look at the results & save them:
SSWresults 

save.image("SSWresults.RData")
```

### Step 7 : Create model 1 for goodness of fit
```{r}
# Included from getEffects()
SSWmodel <- getEffects(SSWdata)

# Add the transitive triplets effects.
model1 <- includeEffects(SSWmodel, transTrip, name="friendship")

# Check what is specified in model1:
model1

# Use the siena07() function to estimate the parameters of model1.
# To use the goodness of fit testing functionality, it is necessary
# to specify returnDeps=TRUE; this will include the simulated networks
# in the results object, and will permit goodness of fit checking later on.
( results1 <- siena07(estimationSettings, data=SSWdata,
                      effects=model1, returnDeps=TRUE) )
```

### Step 8 : GOF using indegreedistribution, outdegreedistribution, geodesicdistribution and triad census
```{r}
# Calculate the fit with with respect to the indegree distribution.
# By specifying "verbose=TRUE" we get information on the screen telling us
# how far calculations have progressed.
gof1.id <- sienaGOF(results1, verbose=TRUE,
                    varName="friendship", IndegreeDistribution)
gof1.id
plot(gof1.id)
# The default for the InDegreeDistribution is to study the fit of the
# cumulative distribution of the indegrees in the network, summed over all waves
# except for the first, for indegrees up to 8.
# The plot shows the observed values as the red dots and numbers;
# e.g., the sum over waves 2 and wave 3 of the numbers of actors with indegree 0
# is 12; the sum of the numbers of actors with indegree at most 3 is 75.
# This applies to the observed data set.
# The rest of the plot refers to the simulated networks in Phase 3
# of the algorithm, where it is supposed to have converged
# (if t-ratios for convergence all are less than 0.10).
# The shapes are so-called violin plots, which combine box plots
# with smoothed density plots.
# The dotted band is a pointwise 90% relative frequency region
# calculated for the simulated data.
# The p-value shows that the indegree distribution is represented well.
# The plot shows the same, as the data are within the band.

# Assess the fit with respect to the outdegree distribution.
gof1.od <- sienaGOF(results1, verbose=TRUE, varName="friendship",
                    OutdegreeDistribution)
gof1.od
plot(gof1.od)
# Here also the p-value and the plot show that the
# outdegree distribution is well represented.

# If the fit of one or both degree distributions had been poor,
# it would have been good to include some degree-related effects
# (outdegree-activity, indegree-popularity, etc.);
# since the fit for the degree distributions is good,
# this is not necessary here.

# Another important set of statistics is the distribution of
# geodesic distances (shortest undirected path lengths between actors).
# GeodesicDistribution calculates the distribution of directed geodesic distances
GeodesicDistribution <- function (i, data, sims, period, groupName,
                                  varName, levls=c(1:5,Inf), cumulative=TRUE, ...) {
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  require(sna)
  a <- sna::geodist(symmetrize(x))$gdist
  if (cumulative)
  {
    gdi <- sapply(levls, function(i){ sum(a<=i) })
  }
  else
  {
    gdi <- sapply(levls, function(i){ sum(a==i) })
  }
  names(gdi) <- as.character(levls)
  gdi
}

# Calculate the fit with respect to the distribution of geodesic distances:
gof1.gd <- sienaGOF(results1, verbose=TRUE,
                    varName="friendship", GeodesicDistribution)
gof1.gd
plot(gof1.gd)
# The fit is very poor, and from the plot we can conclude
# that (except for distances 1, i.e., direct ties)
# there are too many dyads in the simulated networks with small distances.

# Before trying to remedy this, look at the
# fit for the triad census.
TriadCensus <- function(i, data, sims, wave, groupName, varName, levls=1:16){
  unloadNamespace("igraph") # to avoid package clashes
  require(sna)
  require(network)
  x <- networkExtraction(i, data, sims, wave, groupName, varName)
  if (network.edgecount(x) <= 0){x <- symmetrize(x)}
  # because else triad.census(x) will lead to an error
  tc <- sna::triad.census(x)[1,levls]
  # triad names are transferred automatically
  tc
}
gof1.tc <- sienaGOF(results1, verbose=TRUE,
                    varName="friendship", TriadCensus)
# This is more time-consuming...
gof1.tc

# Since the triad counts are wildly different in average and scale,
# plot them using the scale and center options:
plot(gof1.tc, scale=TRUE, center=TRUE)
# The triad census also is unacceptable, although the
# marginal fit across the statistics is not unacceptable.


# Check that the poor fit is not due to time heterogeneity:
tt1 <- sienaTimeTest(results1)
summary(tt1)
# No it isn't.
```

### Step 9 : Create model 2 to improve goodness of fit
```{r}
# Model may be extended to obtain a better fit.
# This is done by a theory-guided exploration.
# In the first place, it is known that the tendency to transitive closure
# might depend less strongly on the number of indirect connections
# than represented by the transitive triplets effect.
# Based on this, a forward model selection is done
# and use the score-type test to see which of these effects
# could improve the model; for the score-type test
# For the gwespFF effect, we use parameter=69;

model2 <- setEffect(model1, cycle3, fix=TRUE, test=TRUE, include=TRUE)
model2 <- setEffect(model2, gwespFF, fix=TRUE, test=TRUE, include=TRUE, parameter=69)
model2 <- setEffect(model2, transTies, fix=TRUE, test=TRUE, include=TRUE)
model2 <- setEffect(model2, transRecTrip, fix=TRUE, test=TRUE, include=TRUE)
model2

( results2 <- siena07(estimationSettings, data=SSWdata,
                      effects=model2, returnDeps=TRUE) )
# The convergence t-ratios will be large for some of the fixed effects;
# this is not a problem at all, because the requirement that they
# should be small applies only to the estimated parameters, not the fixed ones.
# However, these printed results do not contain the score-type tests
# requested by "test=TRUE".
# The joint test has a chi-squared statistic of 16 with d.f.=4,
# (the precise chi-squared value will vary, depending on the randomness
# inherent in the simulation-based procedures),
# which is highly significant.
# Based on this, first include the transitive ties effect.
```

### Step 10 : Create model 3 for goodness of fit
```{r}
# For estimation, use the previous result as the starting value
model3 <- includeEffects(model2, transTies)
( results3 <- siena07(estimationSettings, data=SSWdata,
                      effects=model3, returnDeps=TRUE, prevAns=results2) )
# If any convergence t-ratios for the estimated effects are
# greater than 0.1 in absolute value, estimate again,
# now using prevAns=results3:
( results3 <- siena07(estimationSettings, data=SSWdata,
                    effects=model3, returnDeps=TRUE, prevAns=results3) )
```

### Step 11 : Create model 4 for goodness of fit
```{r}
# Add the transitive reciprocated triplets effect
model4 <- includeEffects(model3, transRecTrip)
( results4 <- siena07(estimationSettings, data=SSWdata,
                      effects=model4, returnDeps=TRUE, prevAns=results3) )
```

### Step 12 : Create model 5 for goodness of fit
```{r}
# The 3-cycle effect still is significant; add it, too:
model5 <- includeEffects(model4, cycle3)
( results5 <- siena07(estimationSettings, data=SSWdata,
                      effects=model5, returnDeps=TRUE, prevAns=results4) )
```

### Step 13 : GOF for model 5
```{r}
# For the new model, apply the same goodness of fit tests.
(gof5.id <- sienaGOF(results5, verbose=TRUE,
                     varName="friendship", IndegreeDistribution))
plot(gof5.id)
# Still good.

(gof5.od <- sienaGOF(results5, verbose=TRUE, varName="friendship",
                     OutdegreeDistribution))
plot(gof5.od)
# Still good.

(gof5.gd <- sienaGOF(results5, verbose=TRUE,
                     varName="friendship", GeodesicDistribution))
plot(gof5.gd)
# There still are too many short distances.
# But the p-value has increased from 0 to 0.001, which shows
# that at least the observation is near the region of simulated values.

(gof5.tc <- sienaGOF(results5, verbose=TRUE,
                     varName="friendship", TriadCensus))
plot(gof5.tc, scale=TRUE, center=TRUE)
# Improved but still not so good.
# Is there now evidence for time heterogeneity?
(tt5 <- sienaTimeTest(results5))
# No.
```

### Step 14 : Create model 6 for goodness of fit
```{r}
# As a next step, the three basic degree-related effects are added.
model6 <- includeEffects(model5, outAct, inPop, outPop)
model6

( results6 <- siena07(estimationSettings, data=SSWdata,
                      effects=model6, returnDeps=TRUE, prevAns=results5) )
# In my case there were some high convergence t-ratios, so I estimated again:
( results6 <- siena07(estimationSettings, data=SSWdata,
                      effects=model6, returnDeps=TRUE, prevAns=results6) )
# Indegree popularity is not significant, so I drop it.
```

### Step 15 : Create model 7 for goodness of fit
```{r}
model7 <- setEffect(model6, inPop, fix=TRUE, test=TRUE, include=TRUE)
( results7 <- siena07(estimationSettings, data=SSWdata,
                      effects=model7, returnDeps=TRUE) )
( results7 <- siena07(estimationSettings, data=SSWdata,
                      effects=model7, returnDeps=TRUE, prevAns=results7) )
```

### Step 16 : GOF for model 7
```{r}
# For the new model, apply the same goodness of fit tests.
(gof7.id <- sienaGOF(results7, verbose=TRUE,
                     varName="friendship", IndegreeDistribution))
plot(gof7.id)
# Still good.

(gof7.od <- sienaGOF(results7, verbose=TRUE, varName="friendship",
                     OutdegreeDistribution))
plot(gof7.od)
# Still good.

(gof7.gd <- sienaGOF(results7, verbose=TRUE,
                     varName="friendship", GeodesicDistribution))
plot(gof7.gd)
# Much improved. There now are moderately too many short distances.

(gof7.tc <- sienaGOF(results7, verbose=TRUE,
                     varName="friendship", TriadCensus))
plot(gof7.tc, scale=TRUE, center=TRUE)
# Good.
# Is there now evidence for time heterogeneity?
(tt7 <- sienaTimeTest(results7))
# No.
```

### Step 17 : Create model 8 for goodness of fit
```{r}
# Check if the gwespFF effect may replace
# the combination of transTrip and transTies;
model8 <- includeEffects(model7, gwespFF)
model8 <- setEffect(model8, transTies, fix=TRUE, test=TRUE, include=TRUE)
model8 <- setEffect(model8, transTrip, fix=TRUE, test=TRUE, include=TRUE)
# First estimate without the prevAns option, because the nonzero estimates
# for transTrip and transTies should not be used:
( results8 <- siena07(estimationSettings, data=SSWdata,
                      effects=model8, returnDeps=TRUE) )
( results8 <- siena07(estimationSettings, data=SSWdata,
                      effects=model8, returnDeps=TRUE, prevAns=results8) )
#summary(results8)
# In this model, transitive triplets and transitive ties
# are not significant any more, so it seems that their
# joint effects are adequately represented by the gwespFF effect.
# Let us check the goodness of fit for this model.
```

### Step 18 : GOF for model 8
```{r}
# For the new model, apply the same goodness of fit tests.
(gof8.id <- sienaGOF(results8, verbose=TRUE,
                     varName="friendship", IndegreeDistribution))
plot(gof8.id)
# Still good.

(gof8.od <- sienaGOF(results8, verbose=TRUE, varName="friendship",
                     OutdegreeDistribution))
plot(gof8.od)
# Still good.

(gof8.gd <- sienaGOF(results8, verbose=TRUE,
                     varName="friendship", GeodesicDistribution))
plot(gof8.gd)
# On the low side but reasonable. There are moderately too many short distances.

(gof8.tc <- sienaGOF(results8, verbose=TRUE,
                     varName="friendship", TriadCensus))
plot(gof8.tc, scale=TRUE, center=TRUE)
# Good.
# Any time heterogeneity?
(tt8 <- sienaTimeTest(results8))
# No.
```

Conclusion : Significant effects are reciprocity, geometrically weighted edgewise shared partners,and 3-cycles (all positive), and transitive reciprocated triplets, outdegree-popularity, and outdegree-activity (all negative). These effects jointly give a good representation of the distributions of indegrees, outdegrees, and geodesic distances,
and of the triad census.











